SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-10 docs/questioning.tex
changes on the exploitation model

  We then categorize and analyze the various defense mechanisms that
  have been developed. Our analysis is based on dimensions such as
  accuracy, performance, deployment, security and availability.
  Especially notable is accuracy, as our findings show that there are
  many defense mechanisms that are tested in a poor manner.
  In addition, we observe that there are protection mechanisms
  that can be bypassed by attackers that know how the mechanisms
  work.

@@

In this section we provide a step-by-step exploitation model to
understand the process of carrying out application attacks based on
source code injection vulnerabilities.
Figure~\ref{fig:attacks} presents the steps of different application
attacks. For every arrow there is a
corresponding tuple that includes the step number for
every attack. For the most part, the steps are common in all attacks. This
is because they are based on similar attack vectors as we mentioned in
the previous section. Most importantly,
Figure~\ref{fig:attacks} also illustrates in which step
the defenses detect corresponding attacks.

A malicious user can initiate the attack through two different routes.
The attacker may use the browser of an innocent user as an attack
vehicle, through which the code will be injected in the application.
For example, the attacker could craft a malicious script into a {\sc url}
and then trick an end user to click on it via social engineering (i.e., phishing ---
{\sc step}: {\sc p-xss}' 1 \text{\textbar} {\sc n-xss}' 1 \text{\textbar} {\sc csrf} 1).
Alternatively, the attacker may be able to inject
directly the malicious code in the application. For example, consider
a web application that accepts and redisplays user input
without appropriate validation. An attacker could
upload data containing a specially crafted script (for example, in a
blog comment) to steal the cookies of the visiting users or embed
malicious {\sc sql} code to retrieve all the password entries from a
database ({\sc step}: {\sc dsl} 1 \text{\textbar}
{\sc p-xss} 1 \text{\textbar} {\sc n-xss} 1).
Note that an {\sc xss} attack can start from both routes.

@@

{\sc xss} attacks fall ino two categories, {\it non-persistent} and
{\it persistent}. Non-persistent {\sc xss} attacks take place when the
data provided by a user is used immediately by server-side scripts to
parse and display a page of results, without properly sanitizing the
request ({\sc step}: {\sc p-xss} 8 \text{\textbar} {\sc n-xss} 3).
Thus, in a non-persistent {\sc xss} attack the injected code
is never saved on the server and immediately becomes part of the
content that is sent back to a web user.

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on source
code injection vulnerabilities. For every arrow there is a
corresponding tuple that includes the step number for
every attack. Note that {\sc xss} attacks can also involve
{\it phishing}. Every {\it point} indicates a step where mechanisms
detect attacks. For every point we provide the corresponding
references:
Point {\sc ub}~\cite{KJKV09,TNH07,NSS06,APKLM10,SPPGMMB04,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,GC09,VFJKKV07,SLMS14,BV08},
Point {\sc sb}~\cite{RDWDE07,JKK06a,JB07,NLC07,WPLKK09,JEP08},
Point {\sc eq}~\cite{BWS05,SW06,PS11,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
Point {\sc sdb}~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (Point {\sc eq}).
% However, this is a simplistic approach that may
% lead to many false positives.
%% PL better to avoid characterisations and mention only the facts
%% (that it leads to many false positives, as shown by ...

It is possible to apply further checks when it is established that
tainted data have reached a sink (Point {\sc eq}). {\it Xu et
  al.}~\cite{XBS06} track taint information at the level of bytes in
memory. To distinguish between legitimate and malicious uses of
untrusted data that reach a sink, they search the data for suspicious
symbols by using regular expressions. {\sc csse}~\cite{PB05}
associates tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application,
and others. When tainted data reaches a sink, {\sc csse} performs
syntactic checks based on the aforementioned metadata (Point {\sc eq}).
{\it {\sc php} Aspis}~\cite{PMP11} works in a similar way. To obtain
metadata, it takes advantage of {\sc php} arrays data structure.
%% PL Like what?
% Note that the above schemes detect potential attacks
% in locus {\sc z}. This automatically means that they
% cannot detect non-persistent JavaScript-driven
% {\sc xss} attacks. Although they can be
% modified to do so.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (Point {\sc sb}).

@@

{\it Vogt et al.}~\cite{VFJKKV07} have developed a tainting scheme
that follows a different approach. Contrary to the aforementioned
schemes that operate on the server side, they track sensitive
information on the client side (Point {\sc ub}). This is a form of
positive taint tracking, where taint data is considered to be
legitimate. Their scheme detects JavaScript-driven {\sc xss} attacks
by ensuring that a script can send sensitive user data only to the
site from which it came from.
% This though, may lead to false alarms.
%% PL You mean a lot of false alarms?
{\it Stock et al.}~\cite{SLMS14} propose a scheme that also operates
in the browser (Point {\sc ub}). The scheme focuses on the detection of
{\sc dom}-based {\sc xss} attacks. In such an attack, the malicious
payload is executed as a result of the modification of the {\sc dom}
environment so that the client-side code runs in an unanticipated
manner even if the {\sc http} response that triggers the injection
seems legitimate. The scheme is different from the previous one
because it marks and observes data that are considered harmful.
Specifically, it employs a taint-enhanced JavaScript engine that
tracks the flow of attacker-controlled data. To detect potential
attacks the scheme utilizes {\sc html} and JavaScript parsers that can
identify the generation of code coming from tainted data.

>>>>

  We then categorize and analyze the various defense mechanisms that
  have been developed. Our analysis is based on dimensions such as
  accuracy, performance, deployment, security and availability.
  Especially notable is accuracy, as our findings show that there are
  many defense mechanisms that are tested in a poor manner.
  In addition, we observe that there are mechanisms
  that can be bypassed by attackers that know how the mechanisms
  work.

@@

In this section we provide a step-by-step exploitation model to
understand the process of carrying out application attacks based on
source code injection vulnerabilities.
Figure~\ref{fig:attacks} presents the steps of different application
attacks. For every arrow there is a
corresponding tuple that includes the transition number for
every attack. For the most part, the steps are common in all attacks. This
is because they are based on similar attack vectors as we mentioned in
the previous section. Most importantly,
Figure~\ref{fig:attacks} also illustrates in which transitions
the defenses detect corresponding attacks.

A malicious user can initiate the attack through two different routes.
The attacker may use the browser of an innocent user as an attack
vehicle, through which the code will be injected in the application.
For example, the attacker could craft a malicious script into a {\sc url}
and then trick an end user to click on it via social engineering (i.e., phishing ---
{\it transition}: {\sc p-xss}' 1 \text{\textbar} {\sc n-xss}' 1 \text{\textbar} {\sc csrf} 1).
Alternatively, the attacker may be able to inject
directly the malicious code in the application. For example, consider
a web application that accepts and redisplays user input
without appropriate validation. An attacker could
upload data containing a specially crafted script (for example, in a
blog comment) to steal the cookies of the visiting users or embed
malicious {\sc sql} code to retrieve all the password entries from a
database ({\it transition}: {\sc dsl} 1 \text{\textbar}
{\sc p-xss} 1 \text{\textbar} {\sc n-xss} 1).
Note that an {\sc xss} attack can start from both routes.

@@

{\sc xss} attacks fall ino two categories, {\it non-persistent} and
{\it persistent}. Non-persistent {\sc xss} attacks take place when the
data provided by a user is used immediately by server-side scripts to
parse and display a page of results, without properly sanitizing the
request ({\sc transition}: {\sc p-xss} 8 \text{\textbar} {\sc n-xss} 3).
Thus, in a non-persistent {\sc xss} attack the injected code
is never saved on the server and immediately becomes part of the
content that is sent back to a web user.

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on source
code injection vulnerabilities. For every arrow there is a
corresponding tuple that includes a transition number for
every attack. Note that {\sc xss} attacks can also involve
phishing. Every {\it point} indicates a transition where mechanisms
detect attacks. For every point we provide the corresponding
references:
Point {\sc ub}~\cite{KJKV09,TNH07,NSS06,APKLM10,SPPGMMB04,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,GC09,VFJKKV07,SLMS14,BV08},
Point {\sc sb}~\cite{RDWDE07,JKK06a,JB07,NLC07,WPLKK09,JEP08},
Point {\sc eq}~\cite{BWS05,SW06,PS11,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
Point {\sc sdb}~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (Point {\sc eq}).
% However, this is a simplistic approach that may
% lead to many false positives.

It is possible to apply further checks when it is established that
tainted data have reached a sink (Point {\sc eq}). {\it Xu et
  al.}~\cite{XBS06} track taint information at the level of bytes in
memory. To distinguish between legitimate and malicious uses of
untrusted data that reach a sink, they search the data for suspicious
symbols by using regular expressions. {\sc csse}~\cite{PB05}
associates tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application,
and others. When tainted data reaches a sink, {\sc csse} performs
syntactic checks based on the aforementioned metadata (Point {\sc eq}).
{\it {\sc php} Aspis}~\cite{PMP11} works in a similar way. To obtain
metadata, it takes advantage of {\sc php} arrays data structure.
% Note that the above schemes detect potential attacks
% in locus {\sc z}. This automatically means that they
% cannot detect non-persistent JavaScript-driven
% {\sc xss} attacks. Although they can be
% modified to do so.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (Point {\sc sb}).

@@

{\it Vogt et al.}~\cite{VFJKKV07} have developed a tainting scheme
that follows a different approach. Contrary to the aforementioned
schemes that operate on the server side, they track sensitive
information on the client side (Point {\sc ub}). This is a form of
positive taint tracking, where taint data is considered to be
legitimate. Their scheme detects JavaScript-driven {\sc xss} attacks
by ensuring that a script can send sensitive user data only to the
site from which it came from.
{\it Stock et al.}~\cite{SLMS14} propose a scheme that also operates
in the browser (Point {\sc ub}). The scheme focuses on the detection of
{\sc dom}-based {\sc xss} attacks. In such an attack, the malicious
payload is executed as a result of the modification of the {\sc dom}
environment so that the client-side code runs in an unanticipated
manner even if the {\sc http} response that triggers the injection
seems legitimate. The scheme is different from the previous one
because it marks and observes data that are considered harmful.
Specifically, it employs a taint-enhanced JavaScript engine that
tracks the flow of attacker-controlled data. To detect potential
attacks the scheme utilizes {\sc html} and JavaScript parsers that can
identify the generation of code coming from tainted data.
